Determinism, Complexity, and Predictability in Computer Performance

### 2014-07-31 results.tex
editing the ARIMA fit paragraph

Note that the \col points (blue {\color{blue}$+$} icons) are clustered
very tightly in the lower left quadrant of the \naive, random-walk,
and LMA plots in Figure~\ref{fig:wpe_vs_mase_all}, but spread out
horizontally in the \arima plot.  This is because of the way the
\arima process builds ARIMA models \cite{autoARIMA}.  If a KPSS test
of the time series in question indicates that it is nonstationary, the
\arima recipe adds an integration term to the model.  This test gives
mixed results in the case of the \col process, flagging five of the 15
trials as stationary and ten as nonstationary.  ARIMA models without
an integration term perform more poorly on this signal, which
increases the error, thereby spreading out the points.  We tested this
hypothesis by forcing the inclusion of an integration term in the five
cases where a KPSS test indicated that such a term was not needed.
This action removed the spread, pushing all 15 of the \col ~ \arima
points in Figure~\ref{fig:wpe_vs_mase_all} into a tight cluster.
\alert{Josh to do: this paragraph needs more, in response to the ref's
  gritch about fitting: the stuff about \arima being an automated
  recipe that can give bad results, and the fact that our technique
  can tell when it does.  Look at the response letter for context and
  ammunition---and also put something in there that calls his
  attention to this paragraph.}

>>>>

Note that the \col points (blue {\color{blue}$+$} icons) are clustered
very tightly in the lower left quadrant of the \naive, random-walk,
and LMA plots in Figure~\ref{fig:wpe_vs_mase_all}, but spread out
horizontally in the \arima plot.  This is because of the way the
\arima process builds ARIMA models \cite{autoARIMA}.  If a KPSS test
of the time series in question indicates that it is nonstationary, the
\arima recipe adds an integration term to the model.  This test gives
mixed results in the case of the \col process, flagging five of the 15
trials as stationary and ten as nonstationary.  ARIMA models without
an integration term perform more poorly on this signal, which
increases the error, thereby spreading out the points.  We tested this
hypothesis by forcing the inclusion of an integration term in the five
cases where a KPSS test indicated that such a term was not needed.
This action removed the spread, pushing all 15 of the \col ~ \arima
points in Figure~\ref{fig:wpe_vs_mase_all} into a tight cluster.
%\alert{Josh to do: this paragraph needs more, in response to the %ref's
%  gritch about fitting: the stuff about \arima being an automated
%  recipe that can give bad results, and the fact that our %technique
%  can tell when it does.  Look at the response letter for context %and
%  ammunition---and also put something in there that calls his
%  attention to this paragraph.}

The discussion in the previous paragraph brings to light an interesting feature of our proposed heuristic. The \arima algorithm that was used to build all of the ARIMA
models in this paper, follows the procedure outlined
in Section V B to choose values for the model parameters.  
As discussed previously, this algorithm employs sophisticated tests, such as the KPSS test, to intelligently traverse and reduce model space and then uses well-known criterion for choosing the most-likely of the remaining models. However, it is quite possible that this procedure can choose a sub-optimal fit. Specifically, if the initial space of models being searched is not broad enough, or if one of the preliminary tests gives an erroneous result, then the best ARIMA model may be excluded or eliminated before likelihood is even examined. This circumstance is hard to detect. Maximal-likelihood  tests, such as AIC used in \arima, can only conclude which of the models provided is ``most likely"---but cannot conclude that \emph{none} of the models are likely. In contrast,  
%Sometimes the choices made by that procedure produce a sub-optimal predictor, however there is nothing to say whether a different model from the ARIMA class would work better. 
inappropriate MASE scores (compared to
what the WPE value might suggest) may suggest an inappropriateness in the order selection and parameter estimation procedure, which would flag to a practitioner to reevaluate the appropriateness of fit. For example, 
%During the fitting process of \arima many subclasses of ARIMA models are eliminated based on various tests, such as the KPSS test. If these tests give erroneous results then entire subclasses of ARIMA models are never even considered. Our heuristic may be able to flag when this has occurred. 
by examining the \arima MASE scores vs. WPE values of \col, we were able to see that, while all signal complexities were the same, some of the signals had much higher MASE values. This suggested to us inappropriate parameter selection for these time series. We then went back and manually assisted the automated fitting procedure, resulting in models with consistent MASE vs WPE score, and improved forecasts.
